---
title: "Paper Figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 80)


### This R chunk will get model trajectories and pre-process other object with
### interesting model outputs
library(tidyr)
library(dplyr)
model_dates <- dat$samples[[1]]$trajectories$date[-1L]
emergence_period <- "Emergence3"

### Immunity levels vs infection ----
cut_off <- 0.95
vam_dates <- vam_data %>% filter(region == "england")
end_date <- sircovid::sircovid_date_as_date(max(model_dates))
start_date <- sircovid::sircovid_date_as_date(min(model_dates))

key_dates <- c(start_date,  as.Date("2020-12-08"),
            min(vam_dates$date[vam_dates$Alpha >= 0], na.rm = TRUE),
            min(vam_dates$date[vam_dates$Alpha >= 0.3], na.rm = TRUE),
            min(vam_dates$date[vam_dates$Alpha >= cut_off], na.rm = TRUE),
            min(vam_dates$date[vam_dates$Delta >= 0], na.rm = TRUE),
            min(vam_dates$date[vam_dates$Delta >= 0.3], na.rm = TRUE),
            min(vam_dates$date[vam_dates$Delta >= cut_off], na.rm = TRUE),
            min(vam_dates$date[vam_dates$Omicron >= 0], na.rm = TRUE),
            min(vam_dates$date[vam_dates$Omicron >= 0.3], na.rm = TRUE),
            min(vam_dates$date[vam_dates$Omicron >= cut_off], na.rm = TRUE),
            end_date)
key_events <- c("Start date", "Vacc rollout",
                "Alpha seed",
                "Alpha ≥50%",
                paste0("Alpha ≥", cut_off * 100, "%"),
                "Delta seed",
                "Delta ≥50%",
                paste0("Delta ≥", cut_off * 100, "%"),
                "Omicron seed",
                "Omicron ≥50%",
                paste0("Omicron ≥", cut_off * 100, "%"),
                "End date")

stopifnot(length(key_dates) == length(key_events))

# Proportion of population not in S compartments
S_vaccinated <- get_model_trajectory(dat, "england", "prop_protected_S_vaccinated", TRUE) %>%
  filter(date %in% key_dates) %>%
  mutate_if(is.numeric, round, 3)

# R that have been vaccinated
R_vaccinated <- get_model_trajectory(dat, "england", "prop_protected_R_vaccinated", TRUE) %>%
  filter(date %in% key_dates) %>%
  mutate_if(is.numeric, round, 3)

# R that are unvaccinated
R_unvaccinated <- get_model_trajectory(dat, "england", "prop_protected_R_unvaccinated", TRUE) %>%
  filter(date %in% key_dates) %>%
  mutate_if(is.numeric, round, 3)

# Output proportion of the population with immunity against infection
prop_immunity <- data.frame(
  date = key_dates,
  events = key_events) %>%
  arrange(date)

prop_immunity$Nat_imm_only <- paste0(
  R_unvaccinated$mean * 100, "% (",
  R_unvaccinated$lb * 100, " - ",
  R_unvaccinated$ub * 100, ")")

prop_immunity$Mixed_imm <- paste0(
  R_vaccinated$mean * 100, "% (",
  R_vaccinated$lb * 100, " - ",
  R_vaccinated$ub * 100, ")")

prop_immunity$Vacc_imm_only <- paste0(
  S_vaccinated$mean * 100, "% (",
  S_vaccinated$lb * 100, " - ",
  S_vaccinated$ub * 100, ")")

prop_immunity$Total_imm <- paste0(
  (R_unvaccinated$mean + R_vaccinated$mean + S_vaccinated$mean) * 100, "% (",
  (R_unvaccinated$lb + R_vaccinated$lb + S_vaccinated$lb) * 100, " - ",
  (R_unvaccinated$ub + R_vaccinated$ub + S_vaccinated$ub) * 100, ")"
  )

#----


### Max/min effective severity by variant during the time each dominated ----

what <- c("ihr", "hfr", "ifr")

# Dates at which we have severity change points
strain_epochs <- as.Date(c("2020-09-17", "2021-03-08", "2021-11-01"))
sev_dates <- as.Date(
  c("2020-04-01", "2020-06-01", "2020-06-16", "2020-10-01", "2020-11-15",
  "2021-02-04", "2021-09-17", "2021-10-17", "2021-11-04", "2022-01-16"))

sev <- lapply(what, function(w)
    get_model_severity(dat$severity, vam_data, strain_epochs, w, by_strain = TRUE))
names(sev) <- toupper(what)

process_sev <- function(x) {
  
  x <- x %>%
    filter(region == "england") %>%
    pivot_longer(!c(date, region, dominant_voc)) %>%
    mutate(strain = case_when(
      grepl("strain_1", name) ~ "Strain 1",
      grepl("strain_2", name) ~ "Strain 2",
      TRUE ~ "Weighted")) %>%
    mutate(name = case_when(
      grepl("mean", name) ~ "mean",
      grepl("lb", name) ~ "lb",
      grepl("ub", name) ~ "ub")) %>%
    pivot_wider(names_from = name, values_from = value) %>%
    mutate(strain = case_when(
      strain == "Strain 1" ~ case_when(
        date < strain_epochs[1] ~ NA_character_,
        date >= strain_epochs[1] & date < strain_epochs[2] ~ "Wildtype",
        date >= strain_epochs[2] & date < strain_epochs[3] ~ "Alpha",
        date >= strain_epochs[3] ~ "Delta"),
      strain == "Strain 2" ~ case_when(
        date < strain_epochs[1] ~ NA_character_,
        date >= strain_epochs[1] & date < strain_epochs[2] ~ "Alpha",
        date >= strain_epochs[2] & date < strain_epochs[3] ~ "Delta",
        date >= strain_epochs[3] ~ "Omicron"),
      TRUE ~ "Weighted"))
}

sev <- purrr::map(sev, process_sev)


#----

```


## Main manuscript results

### Evolving transmissibility and control strategies

The R_0 of SARS-CoV-2 variants in England sequentially increased from 
```{r, echo = FALSE}
x <- switch_levels(get_national_intrinsic_values(dat, "Emergence3"))$r0
x <- lapply(x, round, 1)

print(paste0("Wildtype ", x$Wildtype[1], " (95%CrI ", x$Wildtype[2], " - ", x$Wildtype[3],")"))
print(paste0("Alpha ", x$Alpha[1], " (95%CrI ", x$Alpha[2], " - ", x$Alpha[3],")"))
print(paste0("Delta ", x$Delta[1], " (95%CrI ", x$Delta[2], " - ", x$Delta[3],")"))
print(paste0("Omicron ", x$Omicron[1], " (95%CrI ", x$Omicron[2], " - ", x$Omicron[3],")"))

```


Estimate of population immune by early September 2020 (day Alpha seeded)
```{r, echo = FALSE}
print(prop_immunity[prop_immunity$events == "Alpha seed", c("date", "events",  "Total_imm")])
```


Population-level protection (total only) in December 8, 2020, March 2021 (Delta emerges), and late November 2021 (Omicron emerges)
```{r, echo = FALSE}
print(prop_immunity[prop_immunity$events == "Vacc rollout",
                    c("date", "events",  "Total_imm")])
print(prop_immunity[prop_immunity$events == "Delta seed",
                    c("date", "events",  "Total_imm")])
print(prop_immunity[prop_immunity$events == "Omicron seed",
                    c("date", "events",  "Total_imm")])
```

The replacement of the Delta variant by Omicron was driven by the antigenic divergence and thus immune escape properties of Omicron relative to prior variants with population-level protection against infection decreasing
```{r, echo = FALSE}
print(prop_immunity[prop_immunity$events == "Omicron seed",
                    c("date", "events",  "Total_imm")])
print(prop_immunity[prop_immunity$events == paste0("Omicron ≥", cut_off * 100, "%"),
                    c("date", "events",  "Total_imm")])
```


Immunity at the end of the model fit
```{r, echo = FALSE}
print(prop_immunity[prop_immunity$events == "End date", c(3:6)])
```


### Epidemiological drivers of the severity of COVID-19 waves

The basic severity of the variants in England
```{r, echo = FALSE}
x <- switch_levels(get_national_intrinsic_values(dat, "Emergence3"))
basic_ihr <- lapply(x$IHR, round, 3)
basic_hfr <- lapply(x$HFR, round, 3)
basic_ifr <- lapply(x$IFR, round, 3)
```

IHR
```{r, echo = FALSE}
print(paste0("Wildtype ", basic_ihr$Wildtype[1] * 100, "% (95%CrI ", basic_ihr$Wildtype[2]  * 100, " - ", basic_ihr$Wildtype[3]  * 100,")"))
print(paste0("Alpha ", basic_ihr$Alpha[1] * 100, "% (95%CrI ", basic_ihr$Alpha[2]  * 100, " - ", basic_ihr$Alpha[3]  * 100,")"))
print(paste0("Delta ", basic_ihr$Delta[1] * 100, "% (95%CrI ", basic_ihr$Delta[2]  * 100, " - ", basic_ihr$Delta[3]  * 100,")"))
print(paste0("Omicron ", basic_ihr$Omicron[1] * 100, "% (95%CrI ", basic_ihr$Omicron[2]  * 100, " - ", basic_ihr$Omicron[3]  * 100,")"))
```

HFR
```{r, echo = FALSE}
print(paste0("Wildtype ", basic_hfr$Wildtype[1] * 100, "% (95%CrI ", basic_hfr$Wildtype[2]  * 100, " - ", basic_hfr$Wildtype[3]  * 100,")"))
print(paste0("Alpha ", basic_hfr$Alpha[1] * 100, "% (95%CrI ", basic_hfr$Alpha[2]  * 100, " - ", basic_hfr$Alpha[3]  * 100,")"))
print(paste0("Delta ", basic_hfr$Delta[1] * 100, "% (95%CrI ", basic_hfr$Delta[2]  * 100, " - ", basic_hfr$Delta[3]  * 100,")"))
print(paste0("Omicron ", basic_hfr$Omicron[1] * 100, "% (95%CrI ", basic_hfr$Omicron[2]  * 100, " - ", basic_hfr$Omicron[3]  * 100,")"))
```

IFR
```{r, echo = FALSE}
print(paste0("Wildtype ", basic_ifr$Wildtype[1] * 100, "% (95%CrI ", basic_ifr$Wildtype[2]  * 100, " - ", basic_ifr$Wildtype[3]  * 100,")"))
print(paste0("Alpha ", basic_ifr$Alpha[1] * 100, "% (95%CrI ", basic_ifr$Alpha[2]  * 100, " - ", basic_ifr$Alpha[3]  * 100,")"))
print(paste0("Delta ", basic_ifr$Delta[1] * 100, "% (95%CrI ", basic_ifr$Delta[2]  * 100, " - ", basic_ifr$Delta[3]  * 100,")"))
print(paste0("Omicron ", basic_ifr$Omicron[1] * 100, "% (95%CrI ", basic_ifr$Omicron[2]  * 100, " - ", basic_ifr$Omicron[3]  * 100,")"))
```
The percentage point decrease in the basic IFR between Alpha and Delta
```{r, echo = FALSE}
int_sev_eng <- dat$intrinsic_severity_raw$england
i <- which(int_sev_eng$period == "Emergence3")
j_Alpha <- which(int_sev_eng$variant == "Alpha")
j_Delta <- which(int_sev_eng$variant == "Delta")
ifr_diff <- int_sev_eng$IFR[i, j_Alpha, ] - int_sev_eng$IFR[i, j_Delta, ]

print(paste0(round(mean(ifr_diff), 3) * 100, "% (95%CrI ", round(quantile(ifr_diff, 0.025), 3)  * 100, " - ", round(quantile(ifr_diff, 0.975), 3)  * 100,")"))
```


The basic severity of the virus decreased by changes in immunity and healthcare

Delta's effective severity peaked at much lower values than it would have given its basic severity
```{r, echo = FALSE}
sev_HC_delta <- get_national_intrinsic_values(dat, "Mar 2021")$Delta
sev_HC_delta <- lapply(sev_HC_delta, round, 3)


eff_peak_Delta <- lapply(sev, function (s) {
  s %>% filter(dominant_voc == "Delta", strain == "Delta") %>%
  slice(which.max(mean)) %>%
  select(mean, lb, ub) %>%
  c(., recursive = TRUE) %>%
    unname})


ihr_HC <- unname(c(round((sev_HC_delta$IHR[1] / basic_ihr$Delta[1]), 1),
                   round((sev_HC_delta$IHR[2] / basic_ihr$Delta[1]), 1),
                   round((sev_HC_delta$IHR[3] / basic_ihr$Delta[1]), 1)))
hfr_HC <- unname(c(round(1 - (sev_HC_delta$HFR[1] / basic_hfr$Delta[1]), 3) * 100,
                   round(1 - (sev_HC_delta$HFR[3] / basic_hfr$Delta[1]), 3) * 100,
                   round(1 - (sev_HC_delta$HFR[2] / basic_hfr$Delta[1]), 3) * 100))
ifr_HC <- unname(c(round(1 - (sev_HC_delta$IFR[1] / basic_ifr$Delta[1]), 3) * 100,
                   round(1 - (sev_HC_delta$IFR[3] / basic_ifr$Delta[1]), 3) * 100,
                   round(1 - (sev_HC_delta$IFR[2] / basic_ifr$Delta[1]), 3) * 100))


ihr_EF <- unname(c(round(1 - (eff_peak_Delta$IHR[1] / basic_ihr$Delta[1]), 3) * 100,
                   round(1 - (eff_peak_Delta$IHR[3] / basic_ihr$Delta[1]), 3) * 100,
                   round(1 - (eff_peak_Delta$IHR[2] / basic_ihr$Delta[1]), 3) * 100))
hfr_EF <- unname(c(round(1 - (eff_peak_Delta$HFR[1] / basic_hfr$Delta[1]), 3) * 100,
                   round(1 - (eff_peak_Delta$HFR[3] / basic_hfr$Delta[1]), 3) * 100,
                   round(1 - (eff_peak_Delta$HFR[2] / basic_hfr$Delta[1]), 3) * 100))
ifr_EF <- unname(c(round(1 - (eff_peak_Delta$IFR[1] / basic_ifr$Delta[1]), 3) * 100,
                   round(1 - (eff_peak_Delta$IFR[3] / basic_ifr$Delta[1]), 3) * 100,
                   round(1 - (eff_peak_Delta$IFR[2] / basic_ifr$Delta[1]), 3) * 100))

print("Due to changes in healthcare ecosystem:")
print(paste0("IHR ", ihr_HC[1], " (95%CrI ", ihr_HC[2], " - ", ihr_HC[3],") times higher"))
print(paste0("HFR ", hfr_HC[1], "% (95%CrI ", hfr_HC[2], " - ", hfr_HC[3],") lower"))
print(paste0("IFR ", ifr_HC[1], "% (95%CrI ", ifr_HC[2], " - ", ifr_HC[3],") lower"))

print("Due to effective immunity:")
print(paste0("IHR ", ihr_EF[1], "% (95%CrI ", ihr_EF[2], " - ", ihr_EF[3],") lower"))
print(paste0("HFR ", hfr_EF[1], "% (95%CrI ", hfr_EF[2], " - ", hfr_EF[3],") lower"))
print(paste0("IFR ", ifr_EF[1], "% (95%CrI ", ifr_EF[2], " - ", ifr_EF[3],") lower"))
```

Omicron's effective severity peaked at much lower values than it would have given its basic severity
```{r, echo = FALSE}
sev_HC_Omicron <- get_national_intrinsic_values(dat, "Omicron1")$Omicron
sev_HC_Omicron <- lapply(sev_HC_Omicron, round, 3)


eff_peak_Omicron <- lapply(sev, function (s) {
  s %>% filter(dominant_voc == "Omicron", strain == "Omicron") %>%
  slice(which.max(mean)) %>%
  select(mean, lb, ub) %>%
  c(., recursive = TRUE) %>%
    unname})


ihr_HC <- unname(c(round((sev_HC_Omicron$IHR[1] / basic_ihr$Omicron[1]), 1),
                   round((sev_HC_Omicron$IHR[2] / basic_ihr$Omicron[1]), 1),
                   round((sev_HC_Omicron$IHR[3] / basic_ihr$Omicron[1]), 1)))
hfr_HC <- unname(c(round(1 - (sev_HC_Omicron$HFR[1] / basic_hfr$Omicron[1]), 3) * 100,
                   round(1 - (sev_HC_Omicron$HFR[3] / basic_hfr$Omicron[1]), 3) * 100,
                   round(1 - (sev_HC_Omicron$HFR[2] / basic_hfr$Omicron[1]), 3) * 100))
ifr_HC <- unname(c(round(1 - (sev_HC_Omicron$IFR[1] / basic_ifr$Omicron[1]), 3) * 100,
                   round(1 - (sev_HC_Omicron$IFR[3] / basic_ifr$Omicron[1]), 3) * 100,
                   round(1 - (sev_HC_Omicron$IFR[2] / basic_ifr$Omicron[1]), 3) * 100))


ihr_EF <- unname(c(round(1 - (eff_peak_Omicron$IHR[1] / basic_ihr$Omicron[1]), 3) * 100,
                   round(1 - (eff_peak_Omicron$IHR[3] / basic_ihr$Omicron[1]), 3) * 100,
                   round(1 - (eff_peak_Omicron$IHR[2] / basic_ihr$Omicron[1]), 3) * 100))
hfr_EF <- unname(c(round(1 - (eff_peak_Omicron$HFR[1] / basic_hfr$Omicron[1]), 3) * 100,
                   round(1 - (eff_peak_Omicron$HFR[3] / basic_hfr$Omicron[1]), 3) * 100,
                   round(1 - (eff_peak_Omicron$HFR[2] / basic_hfr$Omicron[1]), 3) * 100))
ifr_EF <- unname(c(round(1 - (eff_peak_Omicron$IFR[1] / basic_ifr$Omicron[1]), 3) * 100,
                   round(1 - (eff_peak_Omicron$IFR[3] / basic_ifr$Omicron[1]), 3) * 100,
                   round(1 - (eff_peak_Omicron$IFR[2] / basic_ifr$Omicron[1]), 3) * 100))

print("Due to changes in healthcare ecosystem:")
print(paste0("IHR ", ihr_HC[1], " (95%CrI ", ihr_HC[2], " - ", ihr_HC[3],") times higher"))
print(paste0("HFR ", hfr_HC[1], "% (95%CrI ", hfr_HC[2], " - ", hfr_HC[3],") lower"))
print(paste0("IFR ", ifr_HC[1], "% (95%CrI ", ifr_HC[2], " - ", ifr_HC[3],") lower"))

print("Due to effective immunity:")
print(paste0("IHR ", ihr_EF[1], "% (95%CrI ", ihr_EF[2], " - ", ihr_EF[3],") lower"))
print(paste0("HFR ", hfr_EF[1], "% (95%CrI ", hfr_EF[2], " - ", hfr_EF[3],") lower"))
print(paste0("IFR ", ifr_EF[1], "% (95%CrI ", ifr_EF[2], " - ", ifr_EF[3],") lower"))
```


## Regression model of HFR ~ occupancy 
```{r, echo = FALSE}
options(width = 300)
print(jtools::summ(sev_winter$linear_model))
```








## Other numbers and tables
### Intrinsic severity and R0 by variant
```{r, echo = FALSE}
states <- dat$samples
steps <- dim(states[[1]]$trajectories$state)[3]
cum_inf <- mean(states$england$trajectories$state["infections", , steps])
cum_adm <- mean(states$england$trajectories$state["admitted", , steps]) +
  mean(states$england$trajectories$state["diagnoses", , steps])


int_sev <- dat$intrinsic_severity %>%
  filter(period == emergence_period) %>%
  filter(region == "england")
  
r0 <- get_r0_region(dat, "england") %>%
  mutate(estimate = value) %>%
  select(!value) %>%
  pivot_longer(!c(estimate, region, source)) %>%
  select(estimate, name, value, region, source)
    
tmp <- int_sev %>%
  select(!period) %>%
  rbind(., r0)

natl_int_sev <- NULL
for (v in unique(tmp$name)) {
  
  out <- NULL
  for (i in unique(tmp$source)) {
    mean <- round(sum((tmp %>% filter(source == i, estimate == "mean", name == v))$value), 4)
    lb <- round(sum((tmp %>% filter(source == i, estimate == "lb", name == v))$value), 4)
    ub <- round(sum((tmp %>% filter(source == i, estimate == "ub", name == v))$value), 4)
    
    if (i == "r0") {
      x <- data.frame(mean = mean, lb = lb, ub = ub)
    } else {
      x <- data.frame(mean = mean * 100, lb = lb * 100, ub = ub * 100)
    }
    
    x <- x %>%
      mutate(CI = paste0("(", lb, " - ", ub, ")")) %>%
      summarise(value = paste0(mean, " ", CI))
    
    out <- rbind(out, x)
  }
  
  out <- as.data.frame(out) %>%
    mutate(measure = unique(tmp$source),
           variant = v) %>%
    select(variant, measure, value)
  
  
  natl_int_sev <- rbind(natl_int_sev, out)
  
}

natl_int_sev <- natl_int_sev %>% pivot_wider(names_from = variant)

rmarkdown::paged_table(natl_int_sev, options = list(rows.print = 16))
```

### Effective severity by variant
```{r, echo = FALSE}
further_process_sev <- function(x) {
  
  x <- x %>%
    filter(strain == "Weighted") %>%
    group_by(dominant_voc)
  
  filter_sev <- function (y) {
    max <- y %>%
      filter(mean == max(mean)) %>%
      mutate(range = "max")
    
    min <- y %>%
      filter(mean == min(mean)) %>%
      mutate(range = "min")
    
    rbind(max, min) %>%
      ungroup() %>%
      arrange(dominant_voc) %>%
      mutate_if(is.numeric,  list(~ round(., 4) * 100)) %>%
      mutate(value = paste0(mean, "% (", lb, " - ", ub, ")")) %>%
      select(!c(date, region, strain, mean, lb, ub))
  }
    
  filter_sev(x)
}

further_sev <- purrr::map(sev, further_process_sev)

further_sev <- purrr::map_df(further_sev, ~as.data.frame(.x), .id = "id") %>%
  pivot_wider(names_from = id)


rmarkdown::paged_table(further_sev, options = list(row.s.print = 8))
```

### Table of proportion protected against infection
```{r, echo = FALSE}
rmarkdown::paged_table(prop_immunity, options = list(cols.print = 6, rows.print = 11))
```

### Demography of hospital admissions and deaths
```{r, echo = FALSE}
age_bands <- get_age_band_labels(TRUE)

demo <- list(admissions = get_model_demography(dat, age_bands, "england", "admissions") ,
             deaths = get_model_demography(dat, age_bands, "england", "deaths_hosp"))

dates <- c(min(demo$deaths$date),
           seq.Date(as.Date("2020-11-01"), as.Date("2021-06-01"), "month"),
           max(demo$deaths$date))

demo_process <- function(x) {
  x %>% 
    pivot_wider() %>%
    mutate(total = rowSums(.[-1L])) %>%
    mutate_if(is.numeric,  list(~ . /total)) %>%
    filter(date %in% dates) %>%
    select(date, `80 plus`) %>%
    mutate(rest = 1 - `80 plus`) %>%
    mutate_if(is.numeric,  list(~ round(., 5) * 100))
}

demo <- purrr::map(demo, demo_process)
```
```{r}
## Admissions
rmarkdown::paged_table(demo$admissions, options = list(rows.print = 10))

## Deaths
rmarkdown::paged_table(demo$deaths, options = list(rows.print = 10))
```


### Age specific IHR and HFR driving demography of hospital admissions and deaths
```{r, echo = FALSE}
eff_sev_age <- list(
  ihr = get_eff_sev_age(dat, "england", "ihr", vam_data, age_bands,
                              age_bands_select = FALSE, vacc_uptake = TRUE,
                              by_class = FALSE, plot = FALSE),
  hfr = get_eff_sev_age(dat, "england", "hfr", vam_data, age_bands,
                              age_bands_select = FALSE, vacc_uptake = TRUE,
                              by_class = FALSE, plot = FALSE))

eff_sev_process <- function(x) {
  x <- x %>%
    filter(region == "england") %>%
    mutate_if(is.numeric, list(~ round(., 5) * 100)) %>%
    select(!c(vacc_class, region)) %>%
    pivot_wider(names_from = estimate, values_from = value) %>%
    group_by(dominant_voc, age)
  
  max <- x %>% 
    filter(mean == max(mean)) %>%
    mutate(id = "max")
  
  min <- x %>% 
    filter(mean == min(mean)) %>%
    mutate(id = "min")
  
  x <- rbind(max, min) %>%
    ungroup() %>%
    group_by(date, dominant_voc, age, id) %>%
    mutate(CI = paste0("(", lb, " - ", ub, ")")) %>%
    summarise(value = paste0(mean, " ", CI)) %>%
    mutate(age = factor(age, levels = seq(0, 80, 5))) %>%
    arrange(age)
  
  x <- split(x, f = x$dominant_voc)
}

eff_sev <- suppressMessages(purrr::map(eff_sev_age, eff_sev_process))
```
```{r}
# IHR Alpha
rmarkdown::paged_table(eff_sev$ihr$Alpha, options = list(rows.print = 8))
# HFR Alpha
rmarkdown::paged_table(eff_sev$hfr$Alpha, options = list(rows.print = 8))
```


## Risk ratio  

```{r, echo = FALSE}

get_region_variants <- function(r) {
    pars <- dat$samples[[r]]$pars
    
    # Take regional Rt_general at the start of the model for Wildtype R0
    # then parametrically calculate for the VOCs
    Wildtype <- dat$rt[[r]]$Rt_general[1, "weighted", ]
    Alpha <- Wildtype * pars[, "ta_alpha"]
    Delta <- Alpha * pars[, "ta_delta"]
    Omicron <- Delta * pars[, "ta_omicron"]
    
    list(Wildtype = Wildtype,
         Alpha = Alpha,
         Delta = Delta,
         Omicron = Omicron)
}

reg <- sircovid::regions("england")
r0_variants <- lapply(reg, get_region_variants)
r0_variants <- spimalot:::list_transpose(r0_variants)
    
get_reg_pop <- function(r) {
  p <- dat$samples[[r]]$predict$transform(dat$samples[[r]]$pars[1, ])
  sum(p[[1]]$pars$population)
}
    
wts <- vapply(reg, get_reg_pop, numeric(1))
    
r0_variants <- lapply(r0_variants, function(x) {
  apply(spimalot:::abind_quiet(x, along = 2), 1, weighted.mean, w = wts)})


get_risk_ratio <- function(var1, var2, what) {
  
  if (what == "R0") {
    rr <- r0_variants[[var2]] / r0_variants[[var1]]
  } else {
    int_sev_raw <- dat$intrinsic_severity_raw$england
    i <- which(int_sev_raw$period == "Emergence3")
    j1 <- which(int_sev_raw$variant == var1)
    j2 <- which(int_sev_raw$variant == var2)
  
    rr <- int_sev_raw[[what]][i, j2, ] / int_sev_raw[[what]][i, j1, ]
  }
  
  if (var1 == var2) {
    out <- "-"
  } else {
    mean_rr <- sprintf('%#.2f', mean(rr)) 
    lb_rr <- sprintf('%#.2f', round(quantile(rr, 0.025), 2))
    ub_rr <- sprintf('%#.2f', round(quantile(rr, 0.975), 2))
    out <- paste0(mean_rr, " (", lb_rr, " - ", ub_rr, ")" )
  }
  
  out
}

strains <- c("Wildtype", "Alpha", "Delta", "Omicron")

get_risk_ratio_table <- function(what) {
  tab <- outer(strains, strains, Vectorize(function(x, y) get_risk_ratio(x, y, what)))
  colnames(tab) <- strains
  rownames(tab) <- strains
  tab
}

rr_R0 <- get_risk_ratio_table("R0")
rr_ihr <- get_risk_ratio_table("IHR")
rr_hfr <- get_risk_ratio_table("HFR")
rr_ifr <- get_risk_ratio_table("IFR")
```


# RR R_0
```{r, echo = FALSE}
print(rr_R0)
```

# RR IHR
```{r, echo = FALSE}
print(rr_ihr)
```

# RR HFR
```{r, echo = FALSE}
print(rr_hfr)
```

# RR IFR
```{r, echo = FALSE}
print(rr_ifr)
```


## Convergence diagnostics

Gelman-Rubin and ESS convergence diagnostics from full chains:

```{r, echo = FALSE}
diagnostics <- get_convergence_diagnostic(dat)
print(diagnostics)
```
